The Types and Driving Forces of Modifications in the Evolution of an API

### 2015-05-18 thesis/paper.tex
categories->types

\subsection{Software Evolution} \label{software_evolution}
%%Categories of maintenance and types of maintenance
Software evolution is a field that studies the application of software maintenance activities, changes in software processes and the resulting, evolved versions of the software. The concept of software maintenance has existed since the 1960s when it was first introduced to the software development community \cite{lientz1980software}. A set of four categories describing different types of software maintenance \cite{lientz1980software} became the basis upon which twelve new types were developed \cite{chapin2001types}. The twelve types of software evolution and software maintenance, as seen in Table \ref{table:expl_types}, describe a software evolution activity that relates to one of three particular areas; the code, the software and the customer-experienced functionality.

@@

%Structure of the interview
The interview was conducted using a semi-structured approach \cite{robson2002real}. Structure to the interview was provided by organising it around the software evolution theories \cite{chapin2001types, lehman90sview}. Investigative and open-ended questions related to each category and law were asked to provide a basis for comparison between the answers and our initial analysis. The interviewee was encouraged to speak freely, even if it required a change in direction or topic. This was allowed in order to fulfil our exploratory goal of the study.

@@

%Codes
To analyse the data each change identified in the spreadsheet was coded based on which module the change occurred in and two other characteristics: a) if the change added, removed or modified functionality, and b) if the change was a cosmetic change or not.
%Concepts
Grouping the codes by module allowed us to form general concepts of change. However, changes that affected the whole API were categorised on their own.
%Categories
These concepts were then grouped into categories based on common patterns between the concepts.
%Theory
These categories were then used to identify the most notable trends in the evolution of the API. These trends were later used as a basis for discussion during the interview.

@@

The remainder of the changes were defined as general trends. These changes were not individual, but were instead larger changes that effected the whole API. Such examples include a shift towards using interfaces to access features, improved maintainability, a different error-handling approach and the inclusion of additional 3\textit{rd} party libraries. With the exception of 3\textit{rd} party libraries, the general trends have secondary changes which help to categories what effect the changes had on the API.

@@

\smallskip \noindent
\textbf{Groomative  }
%Define
Groomative changes are defined to be changes that aim to make the code more maintainable immediately. This should not be confused with preventive changes.
%Data and examples
While none of the change identified were primarily groomative changes, three of them were secondarily categorised as being groomative. This seems to indicate that the groomative nature of a change is often a bonus of other kinds of changes. The move towards using interfaces can be seen as an example of this. While the primary motive of change was to make the API more maintainable in the future, making the API more maintainable immediately was also a reason for the change.
%Conclusion
While not a primary factor for any of the changes, groomative types of change should still be viewed as important for API evolution.

@@

\noindent
\textbf{A majority of the laws apply  } %Reviewed by JW
Evidence was found for the laws of Continuing Change, Self Regulation, Conservation of Familiarity, Continuing Growth and Feedback System to establish that they apply to APIs. The most prominent explanation for this is that a large part of the introduced changes were related to functionality. It is especially interesting to note that the laws of Self Regulation and Conservation of Familiarity accurately match the API explored in this paper.

\smallskip \noindent
\textbf{Two laws apply to a certain extent  } %Reviewed by JW
The laws of Increasing Complexity and Declining Quality were found to apply to APIs, but the extent to which they apply might not be as significant for APIs compared to other types of systems. The reason that this doubt exists is because these motives were secondary to the motives of increased functionality. An alternative explanation for these laws being of lesser significance in APIs could be that the general inertia to modify APIs which exists \cite{google_talk} \cite{henning2007api} \cite{mcdonnell2013empirical} \cite{robbes2012developers} might dominate the need to improve simplicity and quality.

\smallskip \noindent
\textbf{One law is not applicable  } %Reviewed by JW
Strong evidence was found to suggest that the law of Conservation of Organisational Stability does not apply to APIs. Our results not only show that the law is not true for APIs, but that the opposite of what the law states is true for APIs. This is supported by the consensus in literature of the negative effects caused by modifications to APIs \cite{google_talk} \cite{henning2007api} \cite{mcdonnell2013empirical} \cite{robbes2012developers}. The results from our study also revealed that reluctance to change APIs causes the releases of new versions to be few and sporadic.

\smallskip \noindent
\textbf{The importance of a Feedback System  } %Reviewed by JW
Especially interesting results were found in regards to the law of Feedback System. The importance of having a structured and developed feedback loop was stressed during the interview and it was also expressed as one of the main points of improvement. This suggests that including feedback from more sources and using this feedback to improve the evolution process should be a consideration for APIs.

\subsection{RQ3 Results}
RQ3: \textbf{How do APIs ``passively'' evolve?}

The

% After cross-examining our results with the interviewee and their responses, we found that 78\% of our main category predictions matched the actual motive of the changes and that 78\% of our secondary categorisations matched the actual motives.

>>>>

\subsection{Software Evolution} \label{software_evolution}
%%Types of maintenance and types of maintenance
Software evolution is a field that studies the application of software maintenance activities, changes in software processes and the resulting, evolved versions of the software. The concept of software maintenance has existed since the 1960s when it was first introduced to the software development community \cite{lientz1980software}. A set of four categories describing different types of software maintenance \cite{lientz1980software} became the basis upon which twelve new types were developed \cite{chapin2001types}. The twelve types of software evolution and software maintenance, as seen in Table \ref{table:expl_types}, describe a software evolution activity that relates to one of three particular areas; the code, the software and the customer-experienced functionality.

@@

%Structure of the interview
The interview was conducted using a semi-structured approach \cite{robson2002real}. Structure to the interview was provided by organising it around the software evolution theories \cite{chapin2001types, lehman90sview}. Investigative and open-ended questions related to each type and law were asked to provide a basis for comparison between the answers and our initial analysis. The interviewee was encouraged to speak freely, even if it required a change in direction or topic. This was allowed in order to fulfil our exploratory goal of the study.

@@

%Codes
To analyse the data each change identified in the spreadsheet was coded based on which module the change occurred in and two other characteristics: a) if the change added, removed or modified functionality, and b) if the change was a cosmetic change or not.
%Concepts
Grouping the codes by module allowed us to form general concepts of change. However, changes that affected the whole API were classified on their own. %"that affected the whole API" re-word this
%Categories
These concepts were then grouped into categories based on common patterns between the concepts.
%Theory
These categories were then used to identify the most notable trends in the evolution of the API. These trends were later used as a basis for discussion during the interview.

@@

The remainder of the changes were defined as general trends. These changes were not individual, but were instead larger changes that effected the whole API. Such examples include a shift towards using interfaces to access features, improved maintainability, a different error-handling approach and the inclusion of additional 3\textit{rd} party libraries. With the exception of 3\textit{rd} party libraries, the general trends have secondary changes which help to classify what effect the changes had on the API.

@@

\smallskip \noindent
\textbf{Groomative  }
%Define
Groomative changes are defined to be changes that aim to make the code more maintainable immediately. This should not be confused with preventive changes.
%Data and examples
While none of the change identified were primarily groomative changes, three of them were secondarily classified as being groomative. This seems to indicate that the groomative nature of a change is often a bonus of other kinds of changes. The move towards using interfaces can be seen as an example of this. While the primary motive of change was to make the API more maintainable in the future, making the API more maintainable immediately was also a reason for the change.
%Conclusion
While not a primary factor for any of the changes, groomative types of change should still be viewed as important for API evolution.

@@

\noindent
\textbf{A majority of the laws apply  } 
Evidence was found for the laws of Continuing Change, Self Regulation, Conservation of Familiarity, Continuing Growth and Feedback System to establish that they apply to APIs. The most prominent explanation for this is that a large part of the introduced changes were related to functionality. It is especially interesting to note that the laws of Self Regulation and Conservation of Familiarity accurately match the API explored in this paper.

\smallskip \noindent
\textbf{Two laws apply to a certain extent  } 
The laws of Increasing Complexity and Declining Quality were found to apply to APIs, but the extent to which they apply might not be as significant for APIs compared to other types of systems. The reason that this doubt exists is because these motives were secondary to the motives of increased functionality. An alternative explanation for these laws being of lesser significance in APIs could be that the general inertia to modify APIs which exists \cite{google_talk} \cite{henning2007api} \cite{mcdonnell2013empirical} \cite{robbes2012developers} might dominate the need to improve simplicity and quality.

\smallskip \noindent
\textbf{One law is not applicable  } 
Strong evidence was found to suggest that the law of Conservation of Organisational Stability does not apply to APIs. Our results not only show that the law is not true for APIs, but that the opposite of what the law states is true for APIs. This is supported by the consensus in literature of the negative effects caused by modifications to APIs \cite{google_talk} \cite{henning2007api} \cite{mcdonnell2013empirical} \cite{robbes2012developers}. The results from our study also revealed that reluctance to change APIs causes the releases of new versions to be few and sporadic.

\smallskip \noindent
\textbf{The importance of a Feedback System  } 
Especially interesting results were found in regards to the law of Feedback System. The importance of having a structured and developed feedback loop was stressed during the interview and it was also expressed as one of the main points of improvement. This suggests that including feedback from more sources and using this feedback to improve the evolution process should be a consideration for APIs.

\subsection{RQ3 Results}
\noindent
RQ3: \textbf{To what extent can we reverse engineer API change decisions?}
\smallskip

After cross-examining our initial analysis with the interviewee's responses, we found that 85\% of our main type predictions matched the actual motive of the change and that 85\% of our secondary type predictions were correct.
